Baseline centering problem

### 2019-01-14 paper/tangotd.tex

\paragraph{Early version of the above: Learning to navigate.} Define a function $V(s,s')$ that
represents the time spent in $s'$ if starting at $s$ and acting optimally
towards $s'$ (just as if the reward was $\1_{s'}$). It is defined by an iterated expectimax to reach $s'$, and
satisfies the Bellman equation
\begin{equation}
V(s,s')=\1_{s=s'}+\gamma \max_a \E_{s_1|s,a} V(s_1,s')
\end{equation}
and it also satisfies
\begin{equation}
\label{eq:optrightbellman}
V(s,s')=\1_{s=s'}+\gamma \sum_{s_{t-1}} V(s,s_{t-1})\max_a
\E_{s_t|s_{t-1},a}\1_{s_t=s'}
\end{equation}
although it's not clear to me how to exploit the last expression. 
\footnote{On-policy it works, I think, because on-policy we can sample
reverse transitions. But I think in that case $V(s,s')$ has to be defined
as a density wrt $\mu(s')$.\\Should also work if we maintain a low-rank
estimate of $V$ as a sum of Diracs over a few $s'$.}
But I
think we have to work from both ends if we want to represent
compositionality of actions. Indeed with the first equation, different values
of $s'$ are just completely separate Bellman problems; the generalization
abilities of the function approximator for $V$ may help, but only so
much. Each time we observe a transition $s\to s'$, we should be able to
update the probabilities to reach $s'$ based on the probabilities to
reach $s$.

@@

By multiplying \eqref{eq:optrightbellman} by $V$, one finds the
Newton--Bellman equation for the optimal function $V$ seen as a matrix:
\begin{equation}
\label{eq:optrightbellman}
V^2_{s_1s_2}=V_{s_1s_2}+\gamma \sum_{s_3} V_{s_1s_3}\max_a
\E_{s_4|s_3,a} V_{s_4s_2}
\end{equation}
TODO DOUBLE-CHECK
and representing $V$ by its density wrt $\rho$ namely
$V_{s_1s_2}=z(s_1,s_2)\rho(s_2)$ this suggests the Newton method
\begin{equation}
z(s_1,s_2)\gets 2z(s_1,s_2)-\E_{s_3\sim \rho} z(s_1,s_3)z(s_3,s_2)+\gamma
\E_{s_3\sim \rho} z(s_1,s_3)\max_a \E_{s_4\sim P(s_4|s_3,a)} z(s_4,s_2)
\end{equation}
satisfied by the optimal $z$ function.

>>>>

\paragraph{Early version of the above: Learning to navigate.} Define a function $V(s,s')$ that
represents the time spent in $s'$ if starting at $s$ and acting optimally
towards $s'$ (just as if the reward was $\1_{s'}$). It is defined by an iterated expectimax to reach $s'$, and
satisfies the Bellman equation
\begin{equation}
\label{eq:optleftbellman}
V(s,s')=\1_{s=s'}+\gamma \max_a \E_{s_1|s,a} V(s_1,s')
\end{equation}
and it also satisfies
\begin{equation}
\label{eq:optrightbellman}
V(s,s')=\1_{s=s'}+\gamma \sum_{s_{t-1}} V(s,s_{t-1})\max_a
\E_{s_t|s_{t-1},a}\1_{s_t=s'}
\end{equation}
although it's not clear to me how to exploit the last expression. 
\footnote{On-policy it works, I think, because on-policy we can sample
reverse transitions. But I think in that case $V(s,s')$ has to be defined
as a density wrt $\mu(s')$.\\Should also work if we maintain a low-rank
estimate of $V$ as a sum of Diracs over a few $s'$.}
But I
think we have to work from both ends if we want to represent
compositionality of actions. Indeed with the first equation, different values
of $s'$ are just completely separate Bellman problems; the generalization
abilities of the function approximator for $V$ may help, but only so
much. Each time we observe a transition $s\to s'$, we should be able to
update the probabilities to reach $s'$ based on the probabilities to
reach $s$.

@@

By multiplying \eqref{eq:optleftbellman} by $V$ on the left, one finds the
Newton--Bellman equation for the optimal function $V$ seen as a matrix:
\begin{equation}
V^2_{s_1s_2}=V_{s_1s_2}+\gamma \sum_{s_3} V_{s_1s_3}\max_a
\E_{s_4|s_3,a} V_{s_4s_2}
\end{equation}
and representing $V$ by its density wrt $\rho$ namely
$V_{s_1s_2}=z^\ast(s_1,s_2)\rho(s_2)$ this suggests the Newton method
\begin{multline}
z^\ast(s_1,s_2)\gets 2z^\ast(s_1,s_2)-\E_{s_3\sim \rho}\,
z^\ast(s_1,s_3)z^\ast(s_3,s_2)\\
+\gamma
\E_{s_3\sim \rho} \,z^\ast(s_1,s_3)\max_a \E_{s_4\sim P(s_4|s_3,a)} z^\ast(s_4,s_2)
\end{multline}
satisfied by the optimal $z$ function.
